---
title: "Visualization"
author: "Alex Lee"
date: "5/17/2017"
output:
  html_document: default
---

```{r setup}
## Set Workspace
setwd("~/R/Classes/ECOL 8540 Intro/ECOL8540")

## Load Packages
library(ggplot2)
library(plotly)
library(lubridate)

## Load Data
mers <- read.csv('cases.csv')

# Correcting errors in mers data
mers$hospitalized[890] <- c('2015-02-20')
mers <- mers[-471,]

mers$onset2 <- ymd(mers$onset)
mers$hospitalized2 <- ymd(mers$hospitalized)

day0 <- min(na.omit(mers$onset2))

mers$epi.day <- as.numeric(mers$onset2 - day0)
```

## Introduction

### Making a Plot

```{r}
# Making a Plot
ggplot(data = mers) +
  geom_bar(mapping = aes(x = epi.day)) +
  labs(x     = 'Epidemic day', 
       y     = 'Case count', 
       title = 'Global count of MERS cases by date of symptom onset', caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```

### By Country

```{r}
# fill = country
ggplot(data = mers) +
  geom_bar(mapping = aes(x    = epi.day,
                         fill = country)) +
  labs(x     = 'Epidemic day', 
       y     = 'Case count', 
       title = 'Global count of MERS cases by date of symptom onset', caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```

#### Position = Fill

```{r}
# position = fill
ggplot(data = mers, position = fill) +
  geom_bar(mapping = aes(x    = epi.day,
                         fill = country)) +
  labs(x     = 'Epidemic day', 
       y     = 'Case count', 
       title = 'Global count of MERS cases by date of symptom onset', caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```

### Polar Chart

```{r}
# coord_flip and coord_polar
ggplot(data = mers, position = fill) +
  geom_bar(mapping = aes(x    = epi.day,
                         fill = country)) +
  labs(x     = 'Epidemic day', 
       y     = 'Case count', 
       title = 'Global count of MERS cases by date of symptom onset', caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv") +
  coord_flip() +
  coord_polar()
```

## Univariate Plots

### Distribution of of infectious period

```{r}
## UNIVARIATE PLOTS ##
mers$infectious.period <- mers$hospitalized2 - mers$onset2
mers$infectious.period <- as.numeric(mers$infectious.period,
                                     units = "days")
ggplot(data = mers) +
  geom_histogram(aes(x = infectious.period)) +
  labs(x = 'Infectious period', 
       y = 'Frequency', 
       title = 'Distribution of calculated MERS infectious period',
       caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.")
```

### Distribution of of infectious period

Positive values only.

```{r}
# Positive infectious periods only
mers$infectious.period2 <- ifelse(mers$infectious.period < 0,
                                  0,
                                  mers$infectious.period)
ggplot(data = mers) +
  geom_histogram(aes(x = infectious.period2)) +
  labs(x = 'Infectious period', 
       y = 'Frequency', 
       title = 'Distribution of calculated MERS infectious period (positive values only)',
       caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.")
```

### Density Plot

```{r}
# Density Plot
ggplot(data = mers) +
  geom_density(mapping = aes(x = infectious.period2)) +
  labs(x = 'Infectious period', 
       y = 'Frequency', 
       title = 'Probability density for MERS infectious period (positive values only)',
       caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.")
```

### Area Plot

```{r}
# Area Plot
ggplot(data = mers) +
  geom_area(stat = 'bin',
            mapping = aes(x = infectious.period2)) +
  labs(x = 'Infectious period', 
       y = 'Frequency', 
       title = 'Area plot for MERS infectious period (positive values only)',
       caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.")
```

### Dot Plot

```{r}
# Dot Plot
ggplot(data = mers) +
  geom_dotplot(mapping = aes(x = infectious.period2)) +
  labs(x = 'Infectious period', 
       y = 'Frequency', 
       title = 'Area plot for MERS infectious period (positive values only)',
       caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.")
```

### Bar Plot 

```{r}
# Bar Plot
ggplot(data = mers) +
  geom_bar(mapping = aes(x = infectious.period2)) +
  labs(x = 'Infectious period', 
       y = 'Frequency', 
       title = 'Area plot for MERS infectious period (positive values only)',
       caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.")
```

## Bivariate Plots

### Infectious Period Over Time

```{r}
## BIVARIATE PLOTS ##

# Infectious period over the course of the MERS epidemic
ggplot(data = mers) +
  geom_line(mapping = aes(y = infectious.period2, 
                          x = onset2))  +
  labs(x = 'Onset',
       y = 'Infectious Period', 
       title = 'Infectious Period During MERS Epidemic (positive values only)',
       caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.")
```

### Societal Learning?

```{r}
# Societal Learning
ggplot(data = mers) +
  geom_line(mapping = aes(y = infectious.period2, 
                          x = onset2))  +
  geom_smooth(mapping = aes(y = infectious.period2, 
                            x = onset2),
              method = "loess")  +
  labs(x = 'Onset', 
       y = 'Infectious Period', 
       title = 'Infectious Period During MERS Epidemic (positive values only)',
       caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.")
```

### By Country

```{r}
# By Country
ggplot(data = mers) +
  geom_line(mapping = aes(y = infectious.period2, 
                          x = onset2))  +
  geom_smooth(mapping = aes(y = infectious.period2, 
                            x = onset2,
                            fill = country),
              method = "loess")  +
  labs(x = 'Onset', 
       y = 'Infectious Period', 
       title = 'Infectious Period During MERS Epidemic (positive values only)',
       caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.")
```

## Faceting

### Country

```{r}
## FACETING ##
# By Country
ggplot(data = mers, 
       mapping = aes(x=epi.day, y=infectious.period2)) + 
  geom_point(mapping = aes(color=country)) +
  facet_wrap(~ country) +
  scale_y_continuous(limits = c(0, 50)) +
  labs(x='Epidemic day', 
       y = 'Infectious period',
       title='MERS infectious period by country', 
       caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.")
```

### Country & Gender

```{r}
# By Country & Gender
ggplot(data = subset(mers, gender %in% c('M', 'F') & country %in% c('KSA', 'Oman', 'Iran', 'Jordan', 'Qatar', 'South Korea', 'UAE' )),
       mapping = aes(x=epi.day, y=infectious.period2 )) +
  geom_point(mapping = aes(color=country)) +
  facet_grid(gender ~ country) +
  scale_y_continuous(limits = c(0, 50)) +
  labs(x = 'Epidemic day', 
       y = 'Infectious period',
       title = 'MERS infectious period by gender and country', 
       caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.")
```

### Case Fatality Rate

Work in progress, code does not work.

```{r, eval=FALSE, include=FALSE}
# Case Fatality Rate
mers.cfr <- 

ggplot(data = subset(mers, outcome %in% c("fatal", "", "recovered") & country %in% c('KSA', 'Oman', 'Iran', 'Jordan', 'Qatar', 'South Korea', 'UAE')),
       mapping = aes(x = epi.day, y = infectious.period2)) +
  geom_point(mapping = aes(color = country)) +
  facet_wrap(~ country) +
  scale_y_continuous(limits = c(0, 50)) +
  labs(x = 'Epidemic day', 
       y = 'Infectious period',
       title = 'MERS infectious period by gender and country', 
       caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.")
```

### Interactive

For HTML output only.

```{r}
# More
epi.curve <- ggplot(data = mers) +
  geom_bar(mapping = aes(x = epi.day)) +
  labs(x = 'Epidemic day', 
       y = 'Case count', 
       title = 'Global count of MERS cases by date of symptom onset',
       caption = "Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
ggplotly(epi.curve)
```

